Paper 2: survey analysis

### 2024-10-08 Rmd/Paper2.Rmd
Removed comments on MLM and OLS reanalysis

---
title:        "Paper 2: survey analysis"
author:       "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
date:         "2024-09-23"
bibliography: "../data/meta_data/references.bib"
csl:          "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), "../output/paper2.html") )})
---

TODO

Check baseline model. Hvordan skal jeg tolke modellen jeg tester? Hva representerer variablene jeg putter inn i modellen, og hva representerer y-verdien?

MLM: TPACK~factor(studiested) er lik OLS: TPACK~factor(studiested)

MLM: TPACK~1 er ikke lik OLS: TPACK~factor(studiested)

Hva er rett måte å sette opp en baseline model på i OLS? (med lm_robust og clusters)

Skal studiested være med som variabel i likningen (fixed effect), eller bare som en cluster (random effect)?

Beregner jeg individuell TPACK eller studiested TPACK? TPACK for den enkelte PST, eller TPACK for hvert studiested?

Hvis studiested er med som variable, beregner jeg TPACK per studiested, ikke student? Hvis studiested ikke er med som variabel, regner jeg TPACK for den enkelte student?

@@

For the multi-level modeling (MLM) the PSTs are considered level 1 units and the TEIs are level 2. The covariates gender and target grade are related to the level 1 units, while year of study can vary at both levels.

@@

Both the OLS and the MLM are done by introducing the variables sequentially. For the OLS done in paper 2, the level 1 variables (i.e., gender, trinn and semester) are introduced first, then the level 2 variables (i.e., TEIs), and finally, PCK and TK. For the MLM, a baseline model (null-model) is created first, where the TEIs are set as random (i.e., the intercept is allowed to vary among TEIs). Then the level 1 variables are introduced to create the level 1-model. Then PCK and TK are added to the model. Since the order of introducing TEI was altered between the OLS in paper 2 and the MLM, a reanalysis of the OLS was done with the TEIs introduced before the covariates. This could affect how much of the variance was explained by the covariates.

The explained variance (i.e., R2) is calculated for both methods, and can be used to compare the results of the two methods.

The following models are tested:

Model | Paper 2 OLS | Reanalysis OLS | New analysis MLM 
------------- | ------------- | ------------- | ------------- 
Options | 3xAnnet = NA | N/A | N/A
Nullmodell | N/A | TPACK ~ TEIs | TPACK ~ 1
Level 1 | TPACK ~ covariates + TEIs | TPACK ~ TEIs + covariates | TPACK ~ covariates
Level 2 | N/A | N/A | + year of study | TEI
PCK | TPACK ~ covariates + TEIs + PCK | TPACK ~ TEIs + covariates + PCK | TPACK ~ covariates + PCK_sentrert
TK | TPACK ~ covariates + TEIs + PCK + TK | TPACK ~ TEIs + covariates + PCK + TK | TPACK ~ covariates + PCK_s + TK_sentrert

The analysis in paper 2 changed the answers of three respondents from "Other" to NA. This made it impossible to compare the fit for the MLMs since the number of respondents were dissimilar for the models compared when including these covariates. For the reanalyses I therefore kept these values unchanged.

In the MLM the TEIs are added as random variables (i.e., the intercepts are allowed to differ between TEIs), while the TEIs are added as dummy variables in the OLS. This should give similar results.

In the MLM the covariates are added as fixed variables (i.e., the slopes are the same for all PSTs independent of TEI).

The order of the models was also changed in the reanalysis, as the baseline model in the MLM corresponds to having the TEIs as dummy variables. In paper 2 this model was tested after having controlled for the covariates. In the reanalysis this order was switched to correspond to the MLM.

Only the MLM has the option to test for level 2 covariates. I allowed slope of year of study vary between TEIs. The level2 model is therefore not tested. It did not converge.

For the MLM the independendent variables were centred using the grand mean, following @christophersen_introduksjon_2018.

The different methods can be compared by comparing R-squared, and diff R-squared.

>>>>

---
title:        "Paper 2: survey analysis"
author:       "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
date:         "2024-10-08"
bibliography: "../data/meta_data/references.bib"
csl:          "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), "../output/Paper2.html") )})
---

@@

@@
